---
title: "Areal disaggregation of reporting probabilities"
output:
  html_notebook:
    toc: true
    toc_float: true
---
This notebook fits person-level binomial areal-disaggregation models with `inlabru`. The target parameter is probability of reporting a VAWG offence. For VAWG definition and data pre=processing steps refer to the main dissertation text.

# 1. Packages and project paths
```{r}
#install.packages(c("survey", "here"))
```

```{r setup, message=FALSE, warning=FALSE}
library(INLA)
library(inlabru)
library(sf)
library(spdep)
library(dplyr)
library(tidyr)
library(purrr)
library(tibble)
library(readr)
library(ggplot2)
library(patchwork)
library(scales)
library(showtext)
library(sysfonts)
library(survey)
library(here)

font_add_google("Roboto Slab", "roboto_slab")
showtext_auto(TRUE)
map_family <- "roboto_slab"

set.seed(20260805) #default seed for predictions
options(scipen = 999, dplyr.summarise.inform = FALSE) #preference to decimal notation

project_dir <- here()
survey_file <- file.path(project_dir, "data/csew/exp_for_inla_stage/export_for_inla_2.csv") #this is the person-level CSEW file
district_file <- file.path(project_dir,"data/geographies/ltla/export_small_geogr_ew_with_covariates_1.gpkg") # LTLA geometry, covariates and age × employment population counts
graph_file <- file.path(project_dir, "data/geographies/ltla/district_graph.gra")
output_dir <- file.path(project_dir, "figures")
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)

compute_options <- bru_options(
  verbose = FALSE,
  control.predictor = list(compute = TRUE),
  control.compute = list(waic = TRUE, dic = TRUE, cpo = TRUE, config = TRUE
  ))
```

# 2. Import survey, spatial data and Census MRP cells

The survey file contains one row per respondent. The LTLA geopackage contains geometry, area-level predictors and the eight Census-derived age × employment poststratification counts. Age and employment are the only MRP dimensions; the remaining individual characteristics stay in the fitted latent field and are not poststratified.

```{r import-data}
snake_case_value <- function(x) { x <- trimws(as.character(x)); x[x==""] <- NA_character_; x <- tolower(x); x <- gsub("&","and",x); x <- gsub("[^a-z0-9]+","_",x); x <- gsub("^_+|_+$","",x); gsub("_+","_",x) }

harmonise_shared_categories <- function(data) data |> mutate(
  agelong=snake_case_value(agelong),gor=snake_case_value(gor),
  remploya=case_when(snake_case_value(remploya)=="employed"~"employed",snake_case_value(remploya)%in%c("not_employed","unemployed","economically_inactive","unemployed_or_economically_inactive")~"unemployed_or_economically_inactive",TRUE~snake_case_value(remploya)),
  nslivarr=case_when(snake_case_value(nslivarr)%in%c("live_as_couple","lives_as_couple","living_as_couple")~"live_as_couple",snake_case_value(nslivarr)%in%c("not_live_as_couple","does_not_live_as_couple","not_living_as_couple")~"not_live_as_couple",TRUE~snake_case_value(nslivarr)),
  educint=case_when(snake_case_value(educint)%in%c("no_qualification","no_qualifications")~"no_qualifications",snake_case_value(educint)%in%c("has_qualification","has_qualifications","qualified")~"has_qualifications",TRUE~snake_case_value(educint)),
  relig3=case_when(snake_case_value(relig3)=="no_religion"~"no_religion",snake_case_value(relig3)%in%c("christian","christian_religion")~"christian_religion",snake_case_value(relig3)%in%c("non_christian","non_christian_religion")~"non_christian_religion",TRUE~snake_case_value(relig3)),
  illharmONS2=case_when(snake_case_value(illharmONS2)=="disabled"~"disabled",snake_case_value(illharmONS2)%in%c("not_disabled","non_disabled")~"not_disabled",TRUE~snake_case_value(illharmONS2))
)

df <- read_csv(survey_file,show_col_types=FALSE)
districts <- st_read(district_file,quiet=TRUE) |> arrange(LAD23CD)

population_cols <- c("age_16_24_Employed","age_25_34_Employed","age_35_64_Employed","age_65_plus_Employed","age_16_24_Unemployed_or_economically_inactive","age_25_34_Unemployed_or_economically_inactive","age_35_64_Unemployed_or_economically_inactive","age_65_plus_Unemployed_or_economically_inactive")
stopifnot(all(population_cols %in% names(districts)),!anyDuplicated(districts$LAD23CD))

mrp_population <- districts |> st_drop_geometry() |> select(LAD23CD,LAD23NM,gor,all_of(population_cols)) |>
  pivot_longer(all_of(population_cols),names_to=c("agelong","remploya"),names_pattern="^(age_(?:16_24|25_34|35_64|65_plus))_(.+)$",values_to="N") |>
  mutate(LAD23CD=as.character(LAD23CD),gor=snake_case_value(gor),agelong=snake_case_value(agelong),remploya=snake_case_value(remploya),N=as.numeric(N)) |>
  filter(is.finite(N),N>0)
```

# 3. Population cells and MRP weights

Both fitting and poststratification use the same LTLA × age × employment Census cells. `fit_weight` distributes a respondent across candidate LTLAs within their GOR and age × employment stratum; `mrp_weight` combines the eight poststratification cells within each LTLA.

```{r population-weights}
district_weights <- mrp_population |>
  mutate(mrp_strata=interaction(agelong,remploya,sep="_",drop=TRUE)) |>
  group_by(LAD23CD,gor,agelong,remploya,mrp_strata) |> summarise(N=sum(N),.groups="drop") |>
  group_by(LAD23CD) |> mutate(mrp_weight=N/sum(N)) |> ungroup()

fit_weights <- district_weights |> group_by(LAD23CD,gor,mrp_strata) |> summarise(N=sum(N),.groups="drop") |>
  group_by(gor,mrp_strata) |> mutate(fit_weight=N/sum(N)) |> ungroup()

stopifnot(all(is.finite(fit_weights$fit_weight)),all(is.finite(district_weights$mrp_weight)),
  max(abs(fit_weights |> group_by(gor,mrp_strata) |> summarise(x=sum(fit_weight),.groups="drop") |> pull(x)-1))<1e-10,
  max(abs(district_weights |> group_by(LAD23CD) |> summarise(x=sum(mrp_weight),.groups="drop") |> pull(x)-1))<1e-10)
```
# 4. LAD adjacency graph

This reproduces the queen-contiguity graph and the three accepted manual links.
The graph order is used to create a stable integer `spatial_id`.

The initial `poly2nb()` call may warn about islands or disconnected subgraphs before
the accepted manual links are added. The checks below are applied to the final graph.

```{r adjacency-graph}
nb <- poly2nb(st_geometry(districts),row.names=districts$LAD23CD,queen=TRUE,snap=350)

add_nb_link <- function(nb,a,b) {
  ids <- attr(nb,"region.id"); ia <- match(a,ids); ib <- match(b,ids)
  stopifnot(!is.na(ia),!is.na(ib),ia!=ib)
  nb[[ia]] <- as.integer(sort(unique(c(nb[[ia]][nb[[ia]]!=0L],ib))))
  nb[[ib]] <- as.integer(sort(unique(c(nb[[ib]][nb[[ib]]!=0L],ia))))
  attr(nb,"sym") <- TRUE
  nb
}

nb <- nb |>
  add_nb_link("E06000046","E07000088") |>
  add_nb_link("E06000046","E06000045") |>
  add_nb_link("E06000053","E06000052")

graph_components <- n.comp.nb(nb)
stopifnot(all(card(nb)>0),graph_components$nc==1L)

dir.create(dirname(graph_file),recursive=TRUE,showWarnings=FALSE)
nb2INLA(graph_file,nb)
graph_obj <- inla.read.graph(graph_file)
graph_levels <- attr(nb,"region.id")

districts <- districts |> mutate(spatial_id=match(LAD23CD,graph_levels))
fit_weights <- fit_weights |> mutate(spatial_id=match(LAD23CD,graph_levels))
district_weights <- district_weights |> mutate(spatial_id=match(LAD23CD,graph_levels))

stopifnot(
  !anyNA(districts$spatial_id),!anyNA(fit_weights$spatial_id),!anyNA(district_weights$spatial_id),
  identical(as.character(districts$LAD23CD[order(districts$spatial_id)]),as.character(graph_levels)),
  graph_obj$n==nrow(districts)
)

check_link <- function(a,b) {
  ia <- match(a,graph_levels); ib <- match(b,graph_levels)
  !is.na(ia) && !is.na(ib) && ib %in% nb[[ia]] && ia %in% nb[[ib]]
}
stopifnot(
  check_link("E06000046","E07000088"),
  check_link("E06000046","E06000045"),
  check_link("E06000053","E06000052")
)

component_summary <- tibble(LAD23CD=graph_levels,component=graph_components$comp.id) |> count(component,name="n_LADs")
component_summary

coords <- st_coordinates(st_centroid(st_geometry(districts)))
plot(st_geometry(districts),border="grey60",col="white",main="Final district neighbour graph")
plot(nb,coords,add=TRUE,col="red",lwd=1)

c(districts=nrow(districts),graph_nodes=graph_obj$n,
  graph_components=graph_components$nc,isolated=sum(card(nb)==0))
```
# 5. Person-level survey preparation

No cluster aggregation is performed. Each retained row is one person with a binary
outcome. The MRP cell combines age and employment. Area characteristics remain
person-linked contextual covariates, as in the original analysis.

All category values are stored in lower-case snake case, including labels such as
`employed`, `has_qualifications`, `christian_religion`, `disabled`, and
`not_live_as_couple`.

```{r prepare-survey}
factor_vars <- c(
  "agelong","relig3","rnssec3","livharm1a","nslivarr","ethgrp2a","remploya","remploy2a",
  "rural3","indlon","inner","imd","imd_crime","polatt7","educint","ladsupgp","nchil2",
  "managhh2","illharmONS2","oa_sup11","income_group","vawg_official_codes",
  "vawg_sex_force_threats","vawg_domestic","vawg_harasm","reported_discard","gor"
)

df <- df |>
  mutate(across(any_of(factor_vars),snake_case_value)) |>
  harmonise_shared_categories() |>
  mutate(
    educint=case_when(educint %in% c("no_qualification","no_qualifications")~"no_qualifications",
                      !is.na(educint)~"has_qualifications",TRUE~NA_character_),
    police_known_all=suppressWarnings(as.integer(as.character(police_known_all))),
    mrp_strata=interaction(agelong,remploya,sep="_",drop=TRUE)
  ) |>
  mutate(across(any_of(factor_vars),factor),mrp_strata=factor(mrp_strata))

area_vars <- c("rural3","inner","imd","imd_crime")
districts <- districts |> mutate(across(any_of(c("gor",area_vars)),snake_case_value))

reference_categories <- list(
  income_group="middle_incomes",managhh2="unexpected_expenses_problem_to_find",
  agelong="age_16_24",relig3="no_religion",ethgrp2a="white",
  imd="middle_deprivation_deciles",imd_crime="middle_deprivation_deciles",
  educint="no_qualifications",nchil2="no_children_in_household",rnssec3="c1",
  polatt7="confident_in_police",illharmONS2="not_disabled",
  remploya="unemployed_or_economically_inactive",nslivarr="not_live_as_couple",
  inner="not_inner_city",rural3="rural",indlon="london",gor="london",
  oa_sup11="hard_pressed_living",vawg_official_codes="false",
  vawg_sex_force_threats="false",vawg_domestic="false",vawg_harasm="false"
)

set_reference_levels <- function(data,references) {
  for(v in intersect(names(references),names(data))) {
    data[[v]] <- droplevels(factor(data[[v]])); ref <- references[[v]]
    if(ref %in% levels(data[[v]])) data[[v]] <- relevel(data[[v]],ref=ref)
    else warning("Reference level not found for ",v,": ",ref,
                 ". Available levels: ",paste(levels(data[[v]]),collapse=", "))
  }
  data
}

df <- set_reference_levels(df,reference_categories)

shared_category_vars <- c("agelong","remploya","gor")
category_mismatches <- map_dfr(shared_category_vars,\(v)
  tibble(
    variable=v,
    survey_only=paste(setdiff(unique(as.character(df[[v]])),unique(as.character(mrp_population[[v]]))),collapse=", "),
    population_only=paste(setdiff(unique(as.character(mrp_population[[v]])),unique(as.character(df[[v]]))),collapse=", ")
  )
)

category_mismatches
```
# 6. Analysis sample and person-to-LAD expansion

All models use one common complete-case sample so WAIC, DIC, CPO, and calibration
comparisons refer to the same respondents. Every person is expanded to all LADs in
their GOR and MRP cell. The person-level covariates are repeated across those candidate
LAD rows; only `spatial_id` and the population weight vary by candidate LAD.

```{r analysis-sample}
offence_vars <- c("vawg_official_codes","vawg_sex_force_threats","vawg_domestic","vawg_harasm")
individual_vars <- c("relig3","rnssec3","ethgrp2a","polatt7","educint","nchil2","illharmONS2","nslivarr")
area_vars <- c("rural3","inner","imd","imd_crime")

required_vars <- c("police_known_all","gor","mrp_strata","ind_weight",offence_vars,individual_vars,area_vars)

person_data <- df |>
  mutate(ind_weight=as.numeric(ind_weight)) |>
  filter(complete.cases(pick(all_of(required_vars))),
         police_known_all %in% c(0L,1L),is.finite(ind_weight),ind_weight>0) |>
  mutate(obs_id=row_number(),y=police_known_all,n=1L)

survey_keys <- person_data |> distinct(gor,mrp_strata) |> mutate(across(everything(),as.character))
population_keys <- fit_weights |> distinct(gor,mrp_strata) |> mutate(across(everything(),as.character))
unmatched_keys <- anti_join(survey_keys,population_keys,by=c("gor","mrp_strata"))

if(nrow(unmatched_keys)) {
  print(unmatched_keys)
  stop(nrow(unmatched_keys)," survey gor × mrp_strata combinations have no population weights.")
}

agg_data <- person_data |>
  left_join(
    fit_weights |> select(gor,mrp_strata,LAD23CD,spatial_id,fit_weight),
    by=c("gor","mrp_strata"),relationship="many-to-many"
  ) |>
  mutate(block=obs_id) |>
  arrange(block,spatial_id)

stopifnot(
  !anyNA(agg_data$spatial_id),!anyNA(agg_data$fit_weight),
  n_distinct(agg_data$block)==nrow(person_data),
  max(abs(agg_data |> group_by(block) |> summarise(x=sum(fit_weight),.groups="drop") |> pull(x)-1))<1e-10
)

c(
  people=nrow(person_data),
  expanded_rows=nrow(agg_data),
  LADs=n_distinct(agg_data$spatial_id),
  GORs=n_distinct(person_data$gor)
)
```

# 7. Survey-weighted direct GOR estimates

The direct estimates use the respondent weight `ind_weight`. They describe observed
reporting differences between GORs and provide an observed-scale comparison for the
model results. They do not validate the inferred LAD estimates.

Because respondents have unequal survey weights, ordinary unweighted Wilson intervals
are not used. Confidence intervals are calculated using the respondent-weighted survey
design.

```{r direct-gor-estimates-1}
# Keep respondents with:
# 1. a valid binary reporting outcome;
# 2. a known GOR;
# 3. a valid positive respondent weight.
direct_data <- person_data |>
  mutate(
    gor = as.character(gor),
    ind_weight = as.numeric(ind_weight)
  ) |>
  filter(
    y %in% c(0L, 1L),
    !is.na(gor),
    is.finite(ind_weight),
    ind_weight > 0
  )


# Create the survey-design object.
#
# ids = ~1 means that only the respondent weights are being used here.
# No additional clustering or stratification variables are supplied.
direct_design <- survey::svydesign(
  ids = ~1,
  weights = ~ind_weight,
  data = direct_data
)


# Calculate the weighted reporting proportion separately for each GOR.
#
# svyciprop() calculates a weighted proportion.
# method = "logit" produces confidence limits that stay between 0 and 1.
direct_gor <- survey::svyby(
  formula = ~y,
  by = ~gor,
  design = direct_design,
  FUN = survey::svyciprop,
  method = "logit",
  vartype = c("se", "ci"),
  level = 0.95,
  na.rm = TRUE,
  keep.names = FALSE
) |>
  as_tibble()


# The survey package normally returns:
# y     = weighted estimated proportion;
# se    = standard error;
# ci_l  = lower confidence limit;
# ci_u  = upper confidence limit.
#
# Rename these columns to make their meaning clearer.
direct_gor <- direct_gor |>
  transmute(
    gor = as.character(gor),
    estimate = as.numeric(y),
    se = as.numeric(`se.as.numeric(y)`),
    lower = as.numeric(ci_l),
    upper = as.numeric(ci_u)
  ) |>

  # Add information about the number of respondents
  # and the total respondent weight in each GOR.
  left_join(
    direct_data |>
      group_by(gor) |>
      summarise(
        unweighted_n = n(),
        weighted_n = sum(ind_weight),
        weighted_reported = sum(ind_weight * y),
        .groups = "drop"
      ),
    by = "gor",
    relationship = "one-to-one"
  ) |>
  arrange(estimate)

names(direct_gor)
# Display the direct weighted estimates.
direct_gor
```


```{r direct-gor-estimates-2, fig.width=6, fig.height=4, dpi=150, out.width="100%"}
# Plot the weighted reporting estimate and its 95% confidence interval
# for each Government Office Region.
p_direct_gor <- ggplot(
  direct_gor,
  aes(
    x = estimate,
    y = reorder(gor, estimate)
  )
) +
  geom_errorbarh(
    aes(
      xmin = lower,
      xmax = upper
    ),
    height = 0.15
  ) +
  geom_point(size = 2) +
  scale_x_continuous(
    limits = c(0, 1),
    labels = label_percent()
  ) +
  labs(
    x = "Survey-weighted reporting estimate with 95% confidence interval",
    y = NULL,
    title = "Observed reporting differences between GORs",
    subtitle = "Person-level estimates weighted by ind_weight"
  ) +
  theme_minimal()

p_direct_gor


# Save the weighted direct estimates.
write_csv(
  direct_gor,
  file.path(
    output_dir,
    "direct_reporting_by_gor_weighted.csv"
  )
)
```

# 8. **PRIORS: primary and sensitivity settings**

**This is the prior-tuning block.** The primary prior follows Wu et al. and the
R-INLA/inlabru reference specification:

- `P(spatial SD > 1) = 0.01`;
- `P(phi > 0.5) = 0.5`.

One alternative is fitted for sensitivity analysis. It more strongly regularises the
total LAD residual variation while leaving the structured/IID mixing prior unchanged:

- `P(spatial SD > 0.5) = 0.05`.

Interpret both SD limits on the log-odds scale and compare the resulting district
surfaces rather than selecting a prior only because it produces a preferred map.

```{r priors}

# ============================ PRIOR BLOCK ============================

prior_settings <- list(

  # Primary prior used in the main analysis.
  "Primary: Wu/reference" = list(
    prec = list(
      prior = "pc.prec",
      param = c(1.0, 0.01)
    ),
    phi = list(
      prior = "pc",
      param = c(0.5, 0.50)
    )
  ),

  # Alternative prior used to check whether the LAD estimates
  # change substantially under stronger spatial regularisation.
  "Sensitivity: stronger regularisation" = list(
    prec = list(
      prior = "pc.prec",
      param = c(0.25, 0.01)
    ),
    phi = list(
      prior = "pc",
      param = c(0.5, 0.50)
    )
  ),
  # Sensitivity 2:
  # same SD prior as primary, different prior on the
  # structured/unstructured variance allocation.
  "Sensitivity: alternative phi prior" = list(
    prec = list(
      prior = "pc.prec",
      param = c(1.0, 0.01)
    ),
    phi = list(
      prior = "pc",
      param = c(0.25, 0.50)
    )
  )
)

# Use the primary prior when fitting the main models.
primary_prior <- prior_settings[["Primary: Wu/reference"]]

# ====================================================================


# Translate example changes on the log-odds scale into probabilities.
#
# This helps interpret what shifts such as -1, -0.5, 0.5 and 1 mean
# when the starting reporting probability is 10%, 20% or 40%.
prior_interpretation <- crossing(
  baseline = c(0.10, 0.20, 0.40),
  shift = c(-1, -0.5, 0.5, 1)
) |>
  mutate(
    probability = plogis(
      qlogis(baseline) + shift
    )
  )

prior_interpretation
```

# 9. Model definitions

The six original substantive model sets are retained. The component formula,
observation formula, and prediction expression are all generated from the same term
list to prevent accidental inconsistencies.

```{r model-definitions}
model_terms <- list(
  "Model 1" = c("rural3", "imd_crime"),
  "Model 2" = c(
    "rural3", "imd_crime", "relig3", "rnssec3", "ethgrp2a",
    "polatt7", "educint", "nchil2", "illharmONS2", "nslivarr"
  ),
  "Model 3" = c("rural3", "inner", "imd", "imd_crime"),
  "Model 4" = c("rural3", "imd_crime", offence_vars),
  "Model 5" = c(area_vars, individual_vars, offence_vars),
  "Model 6" = c(area_vars, individual_vars)
)

component_call <- function(v) sprintf('%s(%s, model = "factor_contrast")', v, v)

make_components <- function(terms, spatial_prior) {
  rhs <- c(
    "-1", "Intercept(1)",
    vapply(terms, component_call, character(1)),
    paste0(
      "district_effect(spatial_id, model = \"bym2\", graph = graph_file, ",
      "scale.model = TRUE, constr = TRUE, hyper = spatial_prior)"
    ),
    'strata_effect(mrp_strata, model = "factor_contrast")'
  )
  as.formula(paste("~", paste(rhs, collapse = " + ")))
}

make_obs_formula <- function(terms) {
  as.formula(paste(
    "y ~",
    paste(c("Intercept", terms, "district_effect", "strata_effect"), collapse = " + ")
  ))
}

make_prediction_formula <- function(terms) {
  as.formula(paste(
    "~ plogis(",
    paste(c("Intercept", terms, "district_effect", "strata_effect"), collapse = " + "),
    ")"
  ))
}
```

# 10. Fit person-level binomial disaggregation models

Each observed person contributes one Bernoulli/binomial outcome. `agg_data` contains
multiple candidate LAD rows for that person. `bm_logitaverage()` computes the
population-weighted average probability over candidate LADs before the binomial
likelihood is evaluated.

```{r fit-models}
fit_disaggregation_model <- function(terms, spatial_prior = primary_prior) {
  components <- make_components(terms, spatial_prior)
  likelihood <- bru_obs(
    family = "binomial",
    formula = make_obs_formula(terms),
    data = agg_data,
    response_data = person_data,
    Ntrials = n,
    aggregate = bm_logitaverage(n_block = nrow(person_data)),
    aggregate_input = list(
      block = agg_data$block,
      weights = agg_data$fit_weight,
      n_block = nrow(person_data)
    )
  )
  bru(components, likelihood, options = compute_options)
}

fits <- imap(model_terms, function(terms, model_name) {
  message("Fitting ", model_name)
  fit_disaggregation_model(terms)
})

list2env(setNames(fits, paste0("fit_", seq_along(fits))), envir = .GlobalEnv)
fits[["Model 1"]]
```

# 11. Fitted probabilities, residuals, and calibration

For Bernoulli outcomes, Pearson residuals use `p(1-p)`. Calibration bins compare the
mean fitted probability with the observed event proportion.

```{r diagnostics, fig.width=20, fig.height=4.5, dpi=150, out.width="100%"}
showtext::showtext_auto(TRUE)
map_family <- "roboto_slab"

extract_diagnostics <- function(fit,model_name,n.samples=1000L) {
  terms <- model_terms[[model_name]]

  candidate_predictions <- predict(
    fit,newdata=agg_data,formula=make_prediction_formula(terms),
    n.samples=n.samples,seed=123,drop=TRUE
  )

  if(is.data.frame(candidate_predictions)) {
    if(!"mean" %in% names(candidate_predictions))
      stop(model_name,": prediction output has no 'mean' column. Columns are: ",paste(names(candidate_predictions),collapse=", "))
    candidate_fitted <- candidate_predictions$mean
  } else if(is.numeric(candidate_predictions)) {
    candidate_fitted <- candidate_predictions
  } else {
    stop(model_name,": unexpected prediction output class: ",paste(class(candidate_predictions),collapse=", "))
  }

  stopifnot(length(candidate_fitted)==nrow(agg_data),all(is.finite(candidate_fitted)))

  fitted_by_person <- tibble(
    block=agg_data$block,fit_weight=agg_data$fit_weight,candidate_fitted=candidate_fitted
  ) |>
    group_by(block) |>
    summarise(fitted=sum(candidate_fitted*fit_weight),.groups="drop") |>
    arrange(block)

  stopifnot(nrow(fitted_by_person)==nrow(person_data),fitted_by_person$block==person_data$obs_id)

  fitted_values <- fitted_by_person$fitted

  residuals <- person_data |>
    transmute(
      obs_id,gor,y,model=model_name,fitted=fitted_values,
      raw_residual=y-fitted_values,
      pearson_residual=(y-fitted_values)/sqrt(pmax(fitted_values*(1-fitted_values),1e-8))
    )

  calibration <- residuals |>
    mutate(bin=ntile(fitted,10)) |>
    group_by(model,bin) |>
    summarise(
      mean_fitted=mean(fitted,na.rm=TRUE),
      mean_observed=mean(y,na.rm=TRUE),
      observations=n(),.groups="drop"
    ) |>
    arrange(model,mean_fitted) |>
    mutate(
      calibration_gap=mean_observed-mean_fitted,
      absolute_gap=abs(calibration_gap),
      ribbon_lower=pmin(mean_fitted,mean_observed),
      ribbon_upper=pmax(mean_fitted,mean_observed)
    )

  brier_score <- mean((residuals$y-residuals$fitted)^2,na.rm=TRUE)
  fitted_safe <- pmin(pmax(residuals$fitted,1e-6),1-1e-6)
  logit_fitted <- qlogis(fitted_safe)

  calibration_intercept <- unname(coef(glm(residuals$y~1,offset=logit_fitted,family=binomial()))[1])
  calibration_slope <- unname(coef(glm(residuals$y~logit_fitted,family=binomial()))[["logit_fitted"]])

  calibration_summary <- calibration |>
    summarise(
      mean_calibration_error=weighted.mean(calibration_gap,w=observations,na.rm=TRUE),
      ece=weighted.mean(absolute_gap,w=observations,na.rm=TRUE),
      maximum_calibration_error=max(absolute_gap,na.rm=TRUE),
      rms_calibration_error=sqrt(weighted.mean(calibration_gap^2,w=observations,na.rm=TRUE))
    ) |>
    mutate(
      brier_score=brier_score,
      calibration_intercept=calibration_intercept,
      calibration_slope=calibration_slope,
      model=model_name
    ) |>
    select(model,everything())

  list(residuals=residuals,calibration=calibration,summary=calibration_summary)
}

diagnostics <- imap(fits,extract_diagnostics)
residuals_all <- map_dfr(diagnostics,"residuals")
calibration_all <- map_dfr(diagnostics,"calibration")
diagnostic_summary <- map_dfr(diagnostics,"summary")
diagnostic_summary

diagnostic_summary_display <- diagnostic_summary |>
  mutate(
    mean_calibration_error_pp=100*mean_calibration_error,
    ece_pp=100*ece,
    maximum_calibration_error_pp=100*maximum_calibration_error,
    rms_calibration_error_pp=100*rms_calibration_error
  ) |>
  select(model,brier_score,calibration_intercept,calibration_slope,
         mean_calibration_error_pp,ece_pp,maximum_calibration_error_pp,rms_calibration_error_pp)

diagnostic_summary_display

diag_theme <- theme_minimal(base_family=map_family,base_size=9) +
  theme(
    text=element_text(family=map_family),
    plot.title=element_text(family=map_family,size=11,face="bold"),
    plot.subtitle=element_text(family=map_family,size=8.5),
    axis.title=element_text(family=map_family,size=8.5),
    axis.text=element_text(family=map_family,size=7.5),
    strip.text=element_text(family=map_family,size=8,face="bold"),
    legend.text=element_text(family=map_family,size=7.5),
    legend.title=element_text(family=map_family,size=8),
    legend.position="bottom",
    panel.grid.minor=element_blank(),
    panel.grid.major=element_line(linewidth=.2),
    plot.margin=margin(4,5,3,4)
  )
```


```{r diagnostics, fig.width=8, fig.height=4, dpi=300, out.width="100%"}
p_residuals <- ggplot(residuals_all, aes(fitted, raw_residual)) +
  geom_point(alpha=.18,size=.45) +
  geom_hline(yintercept=0,linewidth=.35) +
  geom_smooth(se=FALSE,linewidth=.55) +
  facet_wrap(~model,nrow=1) +
  scale_x_continuous(labels=label_percent()) +
  labs(x="Posterior mean fitted reporting probability",
       y="Raw residual (observed - fitted)") +
  diag_theme +
  theme(aspect.ratio=1.03,legend.position="none")

p_calibration <- ggplot(calibration_all, aes(mean_fitted, mean_observed)) +
  geom_ribbon(aes(ymin=ribbon_lower,ymax=ribbon_upper,group=model),alpha=.18) +
  geom_abline(intercept=0,slope=1,linetype="dashed",colour="grey45",linewidth=.25) +
  geom_line(aes(group=model),linewidth=.8) +
  geom_point(size=2) +
  facet_wrap(~model,nrow=1) +
  coord_equal() +
  scale_x_continuous(labels=label_percent()) +
  scale_y_continuous(labels=label_percent()) +
  labs(x="Mean fitted reporting probability",
       y="Observed reporting proportion") +
  diag_theme +
  theme(legend.position="none")

diagnostic_plots <- p_residuals / p_calibration
diagnostic_plots
ggsave(here("figures", "diagnostic_plots.png"),plot=diagnostic_plots, width=8, height=4, dpi=300, bg="white")

write_csv(diagnostic_summary,file.path(output_dir,"model_diagnostic_summary.csv"))
write_csv(diagnostic_summary_display,file.path(output_dir,"model_diagnostic_summary_percentage_points.csv"))
```
# 12. Model comparison
```{r}
fmt <- \(m,l,u) c(coef=sprintf("%.3f",m), cri=sprintf("%.3f to %.3f",l,u))

model_parameters <- imap_dfr(fits,\(fit,m){
  f <- as.data.frame(fit$summary.fixed) |> rownames_to_column("p")
  h <- as.data.frame(fit$summary.hyperpar) |> rownames_to_column("p")
  i <- f |> filter(grepl("Intercept",p)) |> slice(1)
  pr <- h |> filter(grepl("precision.*district",p,ignore.case=TRUE)) |> slice(1)
  ph <- h |> filter(grepl("phi.*district",p,ignore.case=TRUE)) |> slice(1)

  bind_rows(
    tibble(Parameter="Intercept (log-odds scale)", !!!as.list(fmt(i$mean,i$`0.025quant`,i$`0.975quant`))),
    tibble(Parameter="Intercept (probability, %)",
           coef=sprintf("%.1f%%",100*plogis(i$mean)),
           cri=sprintf("%.1f%%–%.1f%%",100*plogis(i$`0.025quant`),100*plogis(i$`0.975quant`))),
    tibble(Parameter="Spatial precision", !!!as.list(fmt(pr$mean,pr$`0.025quant`,pr$`0.975quant`))),
    tibble(Parameter="Spatial SD", !!!as.list(fmt(1/sqrt(pr$mean),1/sqrt(pr$`0.975quant`),1/sqrt(pr$`0.025quant`)))),
    tibble(Parameter="Spatial phi", !!!as.list(fmt(ph$mean,ph$`0.025quant`,ph$`0.975quant`)))
  ) |> mutate(Model=m)
}) |>
  pivot_wider(
    names_from=Model,
    values_from=c(coef,cri),
    names_glue="{Model} {.value}",
    names_vary="slowest"
  )

model_parameters
write_csv(model_parameters,file.path(output_dir, "model_parameters_final.csv"))
```

WAIC, DIC, and CPO describe fit at the observed person/GOR scale. They do not by
themselves validate the inferred LAD pattern.

```{r model-comparison}
model_comparison_table <- imap_dfr(fits,
  function(fit, model_name) {
    tibble(
      Model = model_name,
      WAIC = if (!is.null(fit$waic$waic)) fit$waic$waic else NA_real_, # WAIC and DIC were already calculated during model fitting because control.compute requested them.
      DIC  = if (!is.null(fit$dic$dic)) fit$dic$dic else NA_real_
    )
  }) |>
  mutate(
    delta_WAIC = WAIC - min(WAIC, na.rm = TRUE),
    delta_DIC  = DIC - min(DIC, na.rm = TRUE)
  ) |>
  arrange(WAIC)
model_comparison_table
write_csv(model_comparison_table,file.path(output_dir, "model_comparison_WAIC_DIC_final.csv"))
```

# 13. Parameter and covariate summaries

Age and employment are fitted jointly through `strata_effect(mrp_strata)` and are the only poststratified characteristics. Religion, qualification, disability, living arrangement and the remaining individual, area and offence variables remain fitted model components but are not poststratified. The LAD-specific BYM2 spatial effect is not included in the covariate table.

```{r parameter-tables}
extract_hyperparameters <- function(fit,model_name) as.data.frame(fit$summary.hyperpar) |> rownames_to_column("parameter") |> transmute(model=model_name,parameter,mean,sd,lower=`0.025quant`,median=`0.5quant`,upper=`0.975quant`)

mrp_covariate_vars <- character(0)
non_mrp_individual_vars <- individual_vars
covariate_order <- c("mrp_strata",non_mrp_individual_vars,area_vars,offence_vars)

extract_covariates <- function(fit,model_name) {
  components <- fit$summary.random; keep <- intersect(c("strata_effect",non_mrp_individual_vars,area_vars,offence_vars),names(components))
  map_dfr(keep,\(component) { x <- as.data.frame(components[[component]]); level_col <- if("ID" %in% names(x)) "ID" else names(x)[1]; tibble(model=model_name,covariate=if(component=="strata_effect") "mrp_strata" else component,level=as.character(x[[level_col]]),estimate=as.numeric(x$mean),lower=as.numeric(x$`0.025quant`),upper=as.numeric(x$`0.975quant`)) })
}

hyperparameter_table <- imap_dfr(fits,extract_hyperparameters)
covariate_long <- imap_dfr(fits,extract_covariates) |> mutate(
  group=case_when(covariate=="mrp_strata"~"MRP age × employment strata",covariate%in%non_mrp_individual_vars~"Individual characteristics",covariate%in%area_vars~"Area characteristics",covariate%in%offence_vars~"Offence characteristics",TRUE~"Other"),
  group=factor(group,levels=c("MRP age × employment strata","Individual characteristics","Area characteristics","Offence characteristics","Other")),
  covariate=factor(covariate,levels=covariate_order),coefficient=sprintf("%.3f",estimate),`95% CrI`=sprintf("%.3f to %.3f",lower,upper))

covariate_table_final <- covariate_long |> select(group,covariate,level,model,coefficient,`95% CrI`) |> distinct() |>
  pivot_wider(names_from=model,values_from=c(coefficient,`95% CrI`),names_glue="{model} {.value}",names_vary="slowest") |>
  arrange(group,covariate,level) |> mutate(covariate=as.character(covariate)) |> rename(`Variable group`=group,Variable=covariate,Category=level)

covariate_table_final
mrp_covariate_table <- covariate_table_final |> filter(`Variable group`=="MRP age × employment strata")
non_mrp_covariate_table <- covariate_table_final |> filter(`Variable group`%in%c("Individual characteristics","Area characteristics","Offence characteristics"))
# mrp_covariate_table; non_mrp_covariate_table

# write_csv(hyperparameter_table,file.path(output_dir,"model_hyperparameters.csv"))
# write_csv(covariate_long,file.path(output_dir,"model_covariate_coefficients_long.csv"))
write_csv(covariate_table_final,file.path(output_dir,"covariate_and_mrp_coefficients_by_model_final.csv"))
# write_csv(mrp_covariate_table,file.path(output_dir,"MRP_covariate_coefficients_by_model.csv"))
# write_csv(non_mrp_covariate_table,file.path(output_dir,"non_MRP_covariate_coefficients_by_model.csv"))
```
# 14. LAD-by-MRP prediction grid

Only age and employment are poststratified. Other predictors remain in the fitted model and are fixed at their declared reference categories for LTLA prediction.

```{r prediction-grid}
area_prediction_data <- districts |> st_drop_geometry() |> select(LAD23CD,spatial_id) |> distinct(LAD23CD,.keep_all=TRUE)
poststratified_vars <- c("agelong","remploya")

prediction_data <- district_weights |> select(LAD23CD,gor,spatial_id,mrp_strata,agelong,remploya,N,mrp_weight) |>
  left_join(area_prediction_data,by=c("LAD23CD","spatial_id"),relationship="many-to-one")

non_poststratified_vars <- setdiff(unique(c(area_vars,individual_vars,offence_vars)),poststratified_vars)
for(v in non_poststratified_vars) { if(is.null(reference_categories[[v]])) stop("No reference category declared for: ",v); prediction_data[[v]] <- reference_categories[[v]] }
for(v in intersect(names(prediction_data),names(agg_data))) if(is.factor(agg_data[[v]])) prediction_data[[v]] <- factor(prediction_data[[v]],levels=levels(agg_data[[v]]))

required_prediction_vars <- unique(c("LAD23CD","gor","spatial_id","mrp_strata","N","mrp_weight",unlist(model_terms)))
missing_prediction_vars <- setdiff(required_prediction_vars,names(prediction_data)); if(length(missing_prediction_vars)) stop("Prediction data is missing: ",paste(missing_prediction_vars,collapse=", "))
factor_prediction_vars <- intersect(required_prediction_vars,names(agg_data)); missing_factor_values <- factor_prediction_vars[vapply(prediction_data[factor_prediction_vars],anyNA,logical(1))]
if(length(missing_factor_values)) stop("Prediction variables contain missing values after matching factor levels: ",paste(missing_factor_values,collapse=", "))
stopifnot(!anyNA(prediction_data$spatial_id),!anyNA(prediction_data$mrp_weight),max(abs(prediction_data |> group_by(LAD23CD) |> summarise(x=sum(mrp_weight),.groups="drop") |> pull(x)-1))<1e-10)
prediction_data
```

# 15. Posterior samples and sample-wise MRP
`generate()` returns posterior draws of each LAD-by-cell probability. LAD-overall MRP estimates are calculated within every posterior draw and only then summarised, which preserves dependence between cells and model components.

```{r posterior-prediction}
n_prediction_samples <- 2000L

summarise_draws <- function(draws) {
  tibble(
    mean = rowMeans(draws),
    sd = apply(draws, 1, sd),
    lower = apply(draws, 1, quantile, probs = 0.025),
    median = apply(draws, 1, median),
    upper = apply(draws, 1, quantile, probs = 0.975)
  )
}

predict_model <- function(fit, terms, model_name, data = prediction_data,
                          n.samples = n_prediction_samples) {
  draws <- generate(
    fit, newdata = data,
    formula = make_prediction_formula(terms),
    n.samples = n.samples,
    seed = 123
  )
  draws <- as.matrix(draws)
  stopifnot(nrow(draws) == nrow(data))

  cell <- bind_cols(
    data |> select(LAD23CD,gor,mrp_strata,agelong,remploya,N,mrp_weight),
    summarise_draws(draws)
  ) |>
    mutate(ci_width = upper - lower, model = model_name)

  district_index <- split(seq_len(nrow(data)), data$LAD23CD)
  district_draws <- do.call(rbind, lapply(district_index, function(idx) {
    colSums(draws[idx, , drop = FALSE] * data$mrp_weight[idx])
  }))

  district <- bind_cols(
    tibble(LAD23CD = names(district_index)),
    summarise_draws(district_draws)
  ) |>
    left_join(districts |> st_drop_geometry() |> select(LAD23CD, gor),
              by = "LAD23CD") |>
    mutate(ci_width = upper - lower, model = model_name)

  list(
  cell = cell,
  district = district,
  cell_draws = draws,
  district_draws = district_draws
)
}

predictions <- imap(fits, function(fit, model_name) {
  message("Predicting ", model_name)
  predict_model(fit, model_terms[[model_name]], model_name)
})

cell_predictions <- map_dfr(predictions, "cell")
district_predictions <- map_dfr(predictions, "district")

# write_csv(cell_predictions,
#           file.path(output_dir, "LAD_by_MRP_cell_predictions.csv"))
# write_csv(district_predictions,
#           file.path(output_dir, "LAD_overall_MRP_predictions.csv"))
```

# 16. Aggregation consistency and GOR comparison
`cell_predictions` contains the Census-derived LTLA × age × employment poststratification cells. Overall LTLA estimates use `mrp_weight`; respondent-to-LTLA fitting uses `fit_weight` within GOR × age × employment strata.

```{r aggregation-check}
# Reconstruct model estimates at the observed GOR scale.
# This uses the same respondents and respondent weights as direct_gor,
# making the comparison internally consistent.
reconstructed_gor <- residuals_all |>
  left_join(
    person_data |>
      select(obs_id, ind_weight),
    by = "obs_id",
    relationship = "many-to-one"
  ) |>
  group_by(model, gor) |>
  summarise(
    reconstructed = weighted.mean(
      fitted,
      w = ind_weight,
      na.rm = TRUE
    ),
    .groups = "drop"
  ) |>
  left_join(
    direct_gor |>
      select(
        gor,
        direct = estimate,
        lower,
        upper
      ),
    by = "gor",
    relationship = "many-to-one"
  )
```


```{r}
reconstructed_gor |>
  filter(
    is.na(direct) |
    is.na(reconstructed) |
    direct < 0 | direct > 1 |
    reconstructed < 0 | reconstructed > 1
  ) |>
  select(model, gor, direct, reconstructed)
```
```{r}
df |>
  count(gor, sort = TRUE)

person_data |>
  count(gor, sort = TRUE)

fit_weights |>
  count(gor, sort = TRUE)
```


```{r aggregation-check-2, fig.width=10, fig.height=3, dpi=300, out.width="100%"}
# Both direct and reconstructed values should be probabilities between zero and one.
stopifnot(
  all(reconstructed_gor$direct >= 0 & reconstructed_gor$direct <= 1),
  all(reconstructed_gor$reconstructed >= 0 &
        reconstructed_gor$reconstructed <= 1)
)

direct_reconstructed_gor <- ggplot(
  reconstructed_gor,
  aes(direct, reconstructed, colour = model, group = model)
) +
  geom_abline(intercept = 0, slope = 1, linetype = 2) +
  geom_line(linewidth = 0.6, alpha = 0.7) +
  geom_point() +
  scale_x_continuous(
    limits = c(0.47, 0.66),
    labels = label_percent()
  ) +
  scale_y_continuous(
    limits = c(0.47, 0.66),
    labels = label_percent()
  ) +
  labs(
    x = "Direct GOR reporting estimate",
    y = "Reconstructed model estimate",
    title = "Direct and reconstructed reporting estimates by GOR"
  ) +
  theme_minimal(
    base_family = "roboto_slab"
  ) +
  theme(
    legend.title = element_blank()
  )
direct_reconstructed_gor
ggsave(here("figures", "direct_reconstructed_gor.png"),plot = direct_reconstructed_gor, width = 10,height = 3,dpi = 300,bg = "white")
```

# 17. Prior sensitivity for the primary substantive model
One model is refitted under the stronger-regularisation prior. Posterior LAD estimates are compared with the primary-prior fit. Large LAD changes alongside similar GOR fits indicate that fine-scale conclusions depend materially on the spatial prior.

```{r prior-sensitivity, fig.width=8.27, fig.height=8.83, dpi=300, out.width="100%"}
showtext::showtext_auto(TRUE)
map_family <- "roboto_slab"

# -------------------- PRIOR SENSITIVITY: REFIT + PLOTS --------------------
sensitivity_model <- "Model 5"
sensitivity_terms <- model_terms[[sensitivity_model]]

message("Refitting ", sensitivity_model, " with stronger spatial regularisation...")
fit_sensitivity_prec <- fit_disaggregation_model(
  terms = sensitivity_terms,
  spatial_prior = prior_settings[["Sensitivity: stronger regularisation"]]
)

message("Refitting ", sensitivity_model, " with alternative phi prior...")
fit_sensitivity_phi <- fit_disaggregation_model(
  terms = sensitivity_terms,
  spatial_prior = prior_settings[["Sensitivity: alternative phi prior"]]
)

build_prior_comparison <- function(primary_fit,alt_fit,alt_name,out_file) {
  out <- bind_rows(
    predict_model(primary_fit,sensitivity_terms,"Primary: Wu/reference")$district |> mutate(prior="Primary: Wu/reference"),
    predict_model(alt_fit,sensitivity_terms,alt_name)$district |> mutate(prior=alt_name)
  ) |>
    select(LAD23CD,prior,mean,sd,lower,median,upper,ci_width) |>
    pivot_wider(names_from=prior,values_from=c(mean,sd,lower,median,upper,ci_width),names_glue="{.value}__{prior}") |>
    mutate(
      mean_change=.data[[paste0("mean__",alt_name)]]-.data[["mean__Primary: Wu/reference"]],
      absolute_change=abs(mean_change),
      relative_to_primary_sd=absolute_change/pmax(.data[["sd__Primary: Wu/reference"]],1e-8)
    ) |>
    left_join(
      districts |> st_drop_geometry() |> select(LAD23CD,LAD23NM,gor) |> distinct(),
      by="LAD23CD",relationship="many-to-one"
    ) |>
    mutate(gor=as.character(gor),LAD_label=paste0(LAD23CD,": ",LAD23NM)) |>
    arrange(desc(absolute_change))

  write_csv(out,file.path(output_dir,out_file))
  out
}
prior_comparison_phi <- build_prior_comparison(
  primary_fit = fits[[sensitivity_model]],
  alt_fit = fit_sensitivity_phi,
  alt_name = "Sensitivity: alternative phi prior",
  out_file = "prior_sensitivity_alternative_phi_LAD_comparison.csv"
)

pretty_gor <- function(x) recode(as.character(x),
  "north_east"="North East","north_west"="North West","yorkshire_and_the_humber"="Yorkshire and the Humber",
  "east_midlands"="East Midlands","west_midlands"="West Midlands","east_of_england"="East of England",
  "london"="London","south_east"="South East","south_west"="South West","wales"="Wales",
  .default=stringr::str_to_title(stringr::str_replace_all(as.character(x),"_"," ")))

prior_comparison_prec <- prior_comparison_prec |> mutate(gor=pretty_gor(gor))
prior_comparison_phi <- prior_comparison_phi |> mutate(gor=pretty_gor(gor))

gor_levels <- c("North East","North West","Yorkshire and the Humber","East Midlands","West Midlands","East of England","London","South East","South West","Wales")
gor_cols <- setNames(c("#0072B2","#D55E00","#009E73","#CC79A7","#E69F00","#56B4E9","#000000","#F0E442","#8C564B","#6A3D9A"),gor_levels)

scatter_vals <- c(
  prior_comparison_prec$`mean__Primary: Wu/reference`,
  prior_comparison_prec$`mean__Sensitivity: stronger regularisation`,
  prior_comparison_phi$`mean__Primary: Wu/reference`,
  prior_comparison_phi$`mean__Sensitivity: alternative phi prior`
)
```


```{r prior-sensitivity-0, fig.width=8.27, fig.height=8.83, dpi=300, out.width="100%"}
scatter_lim <- range(scatter_vals,is.finite(scatter_vals),na.rm=TRUE)
scatter_pad <- diff(scatter_lim)*.025
scatter_lim <- c(.49,.57)

prior_theme <- theme_minimal(base_family=map_family) +
  theme(text=element_text(family=map_family),
        plot.title=element_text(size=10.5,face="bold",hjust=0,margin=margin(b=4)),
        axis.title=element_text(size=8.5),axis.title.y=element_text(margin=margin(r=2)),
        axis.text=element_text(size=7.5),
        legend.title=element_blank(),legend.text=element_text(size=7),
        legend.key.width=unit(3,"mm"),legend.key.height=unit(3,"mm"),
        legend.spacing.x=unit(1,"mm"),legend.spacing.y=unit(0,"mm"),
        legend.margin=margin(0,0,0,0),legend.box.margin=margin(0,0,0,0),
        panel.grid.minor=element_blank(),panel.grid.major=element_line(linewidth=.2),
        plot.margin=margin(4,4,4,2))

make_prior_plots <- function(dat,alt_name,scatter_title,change_title,change_xlab) {
  dat <- dat |> mutate(gor=factor(gor,levels=gor_levels))

  p_scatter <- ggplot(dat,aes(x=.data[["mean__Primary: Wu/reference"]],y=.data[[paste0("mean__",alt_name)]],colour=gor)) +
    geom_abline(intercept=0,slope=1,linetype=2,linewidth=.3,colour="grey45") +
    geom_point(alpha=.85,size=.75) +
    scale_colour_manual(values=gor_cols,drop=FALSE) +
    scale_x_continuous(labels=label_percent(),limits=scatter_lim) +
    scale_y_continuous(labels=label_percent(),limits=scatter_lim) +
    coord_equal() +
    labs(x="Primary-prior LAD mean",y="Sensitivity-prior LAD mean",
         title=stringr::str_wrap(scatter_title,30),colour=NULL) +
    guides(colour=guide_legend(nrow=2,byrow=TRUE,override.aes=list(size=2,alpha=1))) +
    prior_theme

  p_change <- dat |>
    slice_max(absolute_change,n=30,with_ties=FALSE) |>
    mutate(LAD_label=reorder(LAD_label,mean_change)) |>
    ggplot(aes(mean_change,LAD_label)) +
    geom_col(fill="#6a2c91") +
    geom_vline(xintercept=0,linetype=2,linewidth=.3) +
    scale_x_continuous(labels=label_percent()) +
    labs(x=change_xlab,y=NULL,title=stringr::str_wrap(change_title,38)) +
    prior_theme + theme(axis.text.y=element_text(size=6.5))

  list(scatter=p_scatter,change=p_change)
}

prec_plots <- make_prior_plots(prior_comparison_prec,"Sensitivity: stronger regularisation",
  "Sensitivity to stronger spatial regularisation",
  "Thirty LTLAs with the largest change under stronger regularisation",
  "Sensitivity mean minus primary mean")

phi_plots <- make_prior_plots(prior_comparison_phi,"Sensitivity: alternative phi prior",
  "Sensitivity to the BYM2 mixing prior",
  "Thirty LTLAs with the largest change under alternative phi",
  "Sensitivity mean minus primary mean")

prior_sensitivity_summaries <- bind_rows(
  prior_comparison_prec |> summarise(
    sensitivity="Stronger regularisation",median_absolute_change=median(absolute_change,na.rm=TRUE),
    p90_absolute_change=quantile(absolute_change,.9,na.rm=TRUE),maximum_absolute_change=max(absolute_change,na.rm=TRUE),
    rank_correlation=cor(`mean__Primary: Wu/reference`,`mean__Sensitivity: stronger regularisation`,method="spearman",use="complete.obs"),
    proportion_change_gt_2pp=mean(absolute_change>.02,na.rm=TRUE),
    proportion_change_gt_primary_sd=mean(relative_to_primary_sd>1,na.rm=TRUE)),
  prior_comparison_phi |> summarise(
    sensitivity="Alternative phi prior",median_absolute_change=median(absolute_change,na.rm=TRUE),
    p90_absolute_change=quantile(absolute_change,.9,na.rm=TRUE),maximum_absolute_change=max(absolute_change,na.rm=TRUE),
    rank_correlation=cor(`mean__Primary: Wu/reference`,`mean__Sensitivity: alternative phi prior`,method="spearman",use="complete.obs"),
    proportion_change_gt_2pp=mean(absolute_change>.02,na.rm=TRUE),
    proportion_change_gt_primary_sd=mean(relative_to_primary_sd>1,na.rm=TRUE))
)

scatter_row <- (prec_plots$scatter | phi_plots$scatter) + plot_layout(guides="collect")
change_row <- prec_plots$change | phi_plots$change

final_prior_plots <- (scatter_row/change_row) +
  plot_annotation(
    title=paste0("Prior sensitivity for ",sensitivity_model),
    subtitle="Top row: primary-prior versus sensitivity-prior LAD means; bottom row: thirty largest LTLA changes",
    theme=theme(text=element_text(family=map_family),
                plot.title=element_text(family=map_family,size=11.5,face="bold",hjust=0),
                plot.subtitle=element_text(family=map_family,size=8.5,hjust=0))
  ) &
  theme(legend.position="bottom",legend.box="horizontal",
        legend.box.spacing=unit(0,"mm"))

print(final_prior_plots)
prior_sensitivity_summaries
```



```{r prior-sensitivity-1, fig.width=8.27, fig.height=8.83, dpi=300, out.width="100%"}
# -------------------- PRIOR SENSITIVITY: REFIT + PLOTS --------------------
sensitivity_model <- "Model 6"
sensitivity_terms <- model_terms[[sensitivity_model]]

message("Refitting ", sensitivity_model, " with stronger spatial regularisation...")
fit_sensitivity_prec <- fit_disaggregation_model(
  terms = sensitivity_terms,
  spatial_prior = prior_settings[["Sensitivity: stronger regularisation"]]
)

message("Refitting ", sensitivity_model, " with alternative phi prior...")
fit_sensitivity_phi <- fit_disaggregation_model(
  terms = sensitivity_terms,
  spatial_prior = prior_settings[["Sensitivity: alternative phi prior"]]
)

build_prior_comparison <- function(primary_fit,alt_fit,alt_name,out_file) {
  out <- bind_rows(
    predict_model(primary_fit,sensitivity_terms,"Primary: Wu/reference")$district |> mutate(prior="Primary: Wu/reference"),
    predict_model(alt_fit,sensitivity_terms,alt_name)$district |> mutate(prior=alt_name)
  ) |>
    select(LAD23CD,prior,mean,sd,lower,median,upper,ci_width) |>
    pivot_wider(names_from=prior,values_from=c(mean,sd,lower,median,upper,ci_width),names_glue="{.value}__{prior}") |>
    mutate(
      mean_change=.data[[paste0("mean__",alt_name)]]-.data[["mean__Primary: Wu/reference"]],
      absolute_change=abs(mean_change),
      relative_to_primary_sd=absolute_change/pmax(.data[["sd__Primary: Wu/reference"]],1e-8)
    ) |>
    left_join(
      districts |> st_drop_geometry() |> select(LAD23CD,LAD23NM,gor) |> distinct(),
      by="LAD23CD",relationship="many-to-one"
    ) |>
    mutate(gor=as.character(gor),LAD_label=paste0(LAD23CD,": ",LAD23NM)) |>
    arrange(desc(absolute_change))

  write_csv(out,file.path(output_dir,out_file))
  out
}

prior_comparison_phi <- build_prior_comparison(
  primary_fit = fits[[sensitivity_model]],
  alt_fit = fit_sensitivity_phi,
  alt_name = "Sensitivity: alternative phi prior",
  out_file = "prior_sensitivity_alternative_phi_LAD_comparison.csv"
)

pretty_gor <- function(x) recode(as.character(x),
  "north_east"="North East","north_west"="North West","yorkshire_and_the_humber"="Yorkshire and the Humber",
  "east_midlands"="East Midlands","west_midlands"="West Midlands","east_of_england"="East of England",
  "london"="London","south_east"="South East","south_west"="South West","wales"="Wales",
  .default=stringr::str_to_title(stringr::str_replace_all(as.character(x),"_"," ")))

prior_comparison_prec <- prior_comparison_prec |> mutate(gor=pretty_gor(gor))
prior_comparison_phi <- prior_comparison_phi |> mutate(gor=pretty_gor(gor))

gor_levels <- c("North East","North West","Yorkshire and the Humber","East Midlands","West Midlands","East of England","London","South East","South West","Wales")
gor_cols <- setNames(c("#0072B2","#D55E00","#009E73","#CC79A7","#E69F00","#56B4E9","#000000","#F0E442","#8C564B","#6A3D9A"),gor_levels)

scatter_vals <- c(
  prior_comparison_prec$`mean__Primary: Wu/reference`,
  prior_comparison_prec$`mean__Sensitivity: stronger regularisation`,
  prior_comparison_phi$`mean__Primary: Wu/reference`,
  prior_comparison_phi$`mean__Sensitivity: alternative phi prior`
)
```


```{r prior-sensitivity-11, fig.width=8.27, fig.height=8.83, dpi=300, out.width="100%"}
scatter_lim <- range(scatter_vals,is.finite(scatter_vals),na.rm=TRUE)
scatter_pad <- diff(scatter_lim)*.025
scatter_lim <- c(.57,.65)

prior_theme <- theme_minimal(base_family=map_family) +
  theme(text=element_text(family=map_family),
        plot.title=element_text(size=10.5,face="bold",hjust=0,margin=margin(b=4)),
        axis.title=element_text(size=8.5),axis.title.y=element_text(margin=margin(r=2)),
        axis.text=element_text(size=7.5),
        legend.title=element_blank(),legend.text=element_text(size=7),
        legend.key.width=unit(3,"mm"),legend.key.height=unit(3,"mm"),
        legend.spacing.x=unit(1,"mm"),legend.spacing.y=unit(0,"mm"),
        legend.margin=margin(0,0,0,0),legend.box.margin=margin(0,0,0,0),
        panel.grid.minor=element_blank(),panel.grid.major=element_line(linewidth=.2),
        plot.margin=margin(4,4,4,2))

make_prior_plots <- function(dat,alt_name,scatter_title,change_title,change_xlab) {
  dat <- dat |> mutate(gor=factor(gor,levels=gor_levels))

  p_scatter <- ggplot(dat,aes(x=.data[["mean__Primary: Wu/reference"]],y=.data[[paste0("mean__",alt_name)]],colour=gor)) +
    geom_abline(intercept=0,slope=1,linetype=2,linewidth=.3,colour="grey45") +
    geom_point(alpha=.85,size=.75) +
    scale_colour_manual(values=gor_cols,drop=FALSE) +
    scale_x_continuous(labels=label_percent(),limits=scatter_lim) +
    scale_y_continuous(labels=label_percent(),limits=scatter_lim) +
    coord_equal() +
    labs(x="Primary-prior LAD mean",y="Sensitivity-prior LAD mean",
         title=stringr::str_wrap(scatter_title,30),colour=NULL) +
    guides(colour=guide_legend(nrow=2,byrow=TRUE,override.aes=list(size=2,alpha=1))) +
    prior_theme

  p_change <- dat |>
    slice_max(absolute_change,n=30,with_ties=FALSE) |>
    mutate(LAD_label=reorder(LAD_label,mean_change)) |>
    ggplot(aes(mean_change,LAD_label)) +
    geom_col(fill="#6a2c91") +
    geom_vline(xintercept=0,linetype=2,linewidth=.3) +
    scale_x_continuous(labels=label_percent()) +
    labs(x=change_xlab,y=NULL,title=stringr::str_wrap(change_title,38)) +
    prior_theme + theme(axis.text.y=element_text(size=6.5))

  list(scatter=p_scatter,change=p_change)
}

prec_plots <- make_prior_plots(prior_comparison_prec,"Sensitivity: stronger regularisation",
  "Sensitivity to stronger spatial regularisation",
  "Thirty LTLAs with the largest change under stronger regularisation",
  "Sensitivity mean minus primary mean")

phi_plots <- make_prior_plots(prior_comparison_phi,"Sensitivity: alternative phi prior",
  "Sensitivity to the BYM2 mixing prior",
  "Thirty LTLAs with the largest change under alternative phi",
  "Sensitivity mean minus primary mean")

prior_sensitivity_summaries <- bind_rows(
  prior_comparison_prec |> summarise(
    sensitivity="Stronger regularisation",median_absolute_change=median(absolute_change,na.rm=TRUE),
    p90_absolute_change=quantile(absolute_change,.9,na.rm=TRUE),maximum_absolute_change=max(absolute_change,na.rm=TRUE),
    rank_correlation=cor(`mean__Primary: Wu/reference`,`mean__Sensitivity: stronger regularisation`,method="spearman",use="complete.obs"),
    proportion_change_gt_2pp=mean(absolute_change>.02,na.rm=TRUE),
    proportion_change_gt_primary_sd=mean(relative_to_primary_sd>1,na.rm=TRUE)),
  prior_comparison_phi |> summarise(
    sensitivity="Alternative phi prior",median_absolute_change=median(absolute_change,na.rm=TRUE),
    p90_absolute_change=quantile(absolute_change,.9,na.rm=TRUE),maximum_absolute_change=max(absolute_change,na.rm=TRUE),
    rank_correlation=cor(`mean__Primary: Wu/reference`,`mean__Sensitivity: alternative phi prior`,method="spearman",use="complete.obs"),
    proportion_change_gt_2pp=mean(absolute_change>.02,na.rm=TRUE),
    proportion_change_gt_primary_sd=mean(relative_to_primary_sd>1,na.rm=TRUE))
)

scatter_row <- (prec_plots$scatter | phi_plots$scatter) + plot_layout(guides="collect")
change_row <- prec_plots$change | phi_plots$change

final_prior_plots <- (scatter_row/change_row) +
  plot_annotation(
    title=paste0("Prior sensitivity for ",sensitivity_model),
    subtitle="Top row: primary-prior versus sensitivity-prior LAD means; bottom row: thirty largest LTLA changes",
    theme=theme(text=element_text(family=map_family),
                plot.title=element_text(family=map_family,size=11.5,face="bold",hjust=0),
                plot.subtitle=element_text(family=map_family,size=8.5,hjust=0))
  ) &
  theme(legend.position="bottom",legend.box="horizontal",
        legend.box.spacing=unit(0,"mm"))

print(final_prior_plots)
prior_sensitivity_summaries
```

```{r prior-sensitivity-2, fig.width=8.27, fig.height=8.83, dpi=300, out.width="100%"}
# -------------------- PRIOR SENSITIVITY: REFIT + PLOTS --------------------
sensitivity_model <- "Model 4"
sensitivity_terms <- model_terms[[sensitivity_model]]

message("Refitting ", sensitivity_model, " with stronger spatial regularisation...")
fit_sensitivity_prec <- fit_disaggregation_model(
  terms = sensitivity_terms,
  spatial_prior = prior_settings[["Sensitivity: stronger regularisation"]]
)

message("Refitting ", sensitivity_model, " with alternative phi prior...")
fit_sensitivity_phi <- fit_disaggregation_model(
  terms = sensitivity_terms,
  spatial_prior = prior_settings[["Sensitivity: alternative phi prior"]]
)

build_prior_comparison <- function(primary_fit,alt_fit,alt_name,out_file) {
  out <- bind_rows(
    predict_model(primary_fit,sensitivity_terms,"Primary: Wu/reference")$district |> mutate(prior="Primary: Wu/reference"),
    predict_model(alt_fit,sensitivity_terms,alt_name)$district |> mutate(prior=alt_name)
  ) |>
    select(LAD23CD,prior,mean,sd,lower,median,upper,ci_width) |>
    pivot_wider(names_from=prior,values_from=c(mean,sd,lower,median,upper,ci_width),names_glue="{.value}__{prior}") |>
    mutate(
      mean_change=.data[[paste0("mean__",alt_name)]]-.data[["mean__Primary: Wu/reference"]],
      absolute_change=abs(mean_change),
      relative_to_primary_sd=absolute_change/pmax(.data[["sd__Primary: Wu/reference"]],1e-8)
    ) |>
    left_join(
      districts |> st_drop_geometry() |> select(LAD23CD,LAD23NM,gor) |> distinct(),
      by="LAD23CD",relationship="many-to-one"
    ) |>
    mutate(gor=as.character(gor),LAD_label=paste0(LAD23CD,": ",LAD23NM)) |>
    arrange(desc(absolute_change))

  write_csv(out,file.path(output_dir,out_file))
  out
}

prior_comparison_phi <- build_prior_comparison(
  primary_fit = fits[[sensitivity_model]],
  alt_fit = fit_sensitivity_phi,
  alt_name = "Sensitivity: alternative phi prior",
  out_file = "prior_sensitivity_alternative_phi_LAD_comparison.csv"
)

pretty_gor <- function(x) recode(as.character(x),
  "north_east"="North East","north_west"="North West","yorkshire_and_the_humber"="Yorkshire and the Humber",
  "east_midlands"="East Midlands","west_midlands"="West Midlands","east_of_england"="East of England",
  "london"="London","south_east"="South East","south_west"="South West","wales"="Wales",
  .default=stringr::str_to_title(stringr::str_replace_all(as.character(x),"_"," ")))

prior_comparison_prec <- prior_comparison_prec |> mutate(gor=pretty_gor(gor))
prior_comparison_phi <- prior_comparison_phi |> mutate(gor=pretty_gor(gor))

gor_levels <- c("North East","North West","Yorkshire and the Humber","East Midlands","West Midlands","East of England","London","South East","South West","Wales")
gor_cols <- setNames(c("#0072B2","#D55E00","#009E73","#CC79A7","#E69F00","#56B4E9","#000000","#F0E442","#8C564B","#6A3D9A"),gor_levels)

scatter_vals <- c(
  prior_comparison_prec$`mean__Primary: Wu/reference`,
  prior_comparison_prec$`mean__Sensitivity: stronger regularisation`,
  prior_comparison_phi$`mean__Primary: Wu/reference`,
  prior_comparison_phi$`mean__Sensitivity: alternative phi prior`
)
```


```{r prior-sensitivity-22, fig.width=8.27, fig.height=8.83, dpi=300, out.width="100%"}
scatter_lim <- range(scatter_vals,is.finite(scatter_vals),na.rm=TRUE)
scatter_pad <- diff(scatter_lim)*.025
scatter_lim <- c(.43,.53)

prior_theme <- theme_minimal(base_family=map_family) +
  theme(text=element_text(family=map_family),
        plot.title=element_text(size=10.5,face="bold",hjust=0,margin=margin(b=4)),
        axis.title=element_text(size=8.5),axis.title.y=element_text(margin=margin(r=2)),
        axis.text=element_text(size=7.5),
        legend.title=element_blank(),legend.text=element_text(size=7),
        legend.key.width=unit(3,"mm"),legend.key.height=unit(3,"mm"),
        legend.spacing.x=unit(1,"mm"),legend.spacing.y=unit(0,"mm"),
        legend.margin=margin(0,0,0,0),legend.box.margin=margin(0,0,0,0),
        panel.grid.minor=element_blank(),panel.grid.major=element_line(linewidth=.2),
        plot.margin=margin(4,4,4,2))

make_prior_plots <- function(dat,alt_name,scatter_title,change_title,change_xlab) {
  dat <- dat |> mutate(gor=factor(gor,levels=gor_levels))

  p_scatter <- ggplot(dat,aes(x=.data[["mean__Primary: Wu/reference"]],y=.data[[paste0("mean__",alt_name)]],colour=gor)) +
    geom_abline(intercept=0,slope=1,linetype=2,linewidth=.3,colour="grey45") +
    geom_point(alpha=.85,size=.75) +
    scale_colour_manual(values=gor_cols,drop=FALSE) +
    scale_x_continuous(labels=label_percent(),limits=scatter_lim) +
    scale_y_continuous(labels=label_percent(),limits=scatter_lim) +
    coord_equal() +
    labs(x="Primary-prior LAD mean",y="Sensitivity-prior LAD mean",
         title=stringr::str_wrap(scatter_title,30),colour=NULL) +
    guides(colour=guide_legend(nrow=2,byrow=TRUE,override.aes=list(size=2,alpha=1))) +
    prior_theme

  p_change <- dat |>
    slice_max(absolute_change,n=30,with_ties=FALSE) |>
    mutate(LAD_label=reorder(LAD_label,mean_change)) |>
    ggplot(aes(mean_change,LAD_label)) +
    geom_col(fill="#6a2c91") +
    geom_vline(xintercept=0,linetype=2,linewidth=.3) +
    scale_x_continuous(labels=label_percent()) +
    labs(x=change_xlab,y=NULL,title=stringr::str_wrap(change_title,38)) +
    prior_theme + theme(axis.text.y=element_text(size=6.5))

  list(scatter=p_scatter,change=p_change)
}

prec_plots <- make_prior_plots(prior_comparison_prec,"Sensitivity: stronger regularisation",
  "Sensitivity to stronger spatial regularisation",
  "Thirty LTLAs with the largest change under stronger regularisation",
  "Sensitivity mean minus primary mean")

phi_plots <- make_prior_plots(prior_comparison_phi,"Sensitivity: alternative phi prior",
  "Sensitivity to the BYM2 mixing prior",
  "Thirty LTLAs with the largest change under alternative phi",
  "Sensitivity mean minus primary mean")

prior_sensitivity_summaries <- bind_rows(
  prior_comparison_prec |> summarise(
    sensitivity="Stronger regularisation",median_absolute_change=median(absolute_change,na.rm=TRUE),
    p90_absolute_change=quantile(absolute_change,.9,na.rm=TRUE),maximum_absolute_change=max(absolute_change,na.rm=TRUE),
    rank_correlation=cor(`mean__Primary: Wu/reference`,`mean__Sensitivity: stronger regularisation`,method="spearman",use="complete.obs"),
    proportion_change_gt_2pp=mean(absolute_change>.02,na.rm=TRUE),
    proportion_change_gt_primary_sd=mean(relative_to_primary_sd>1,na.rm=TRUE)),
  prior_comparison_phi |> summarise(
    sensitivity="Alternative phi prior",median_absolute_change=median(absolute_change,na.rm=TRUE),
    p90_absolute_change=quantile(absolute_change,.9,na.rm=TRUE),maximum_absolute_change=max(absolute_change,na.rm=TRUE),
    rank_correlation=cor(`mean__Primary: Wu/reference`,`mean__Sensitivity: alternative phi prior`,method="spearman",use="complete.obs"),
    proportion_change_gt_2pp=mean(absolute_change>.02,na.rm=TRUE),
    proportion_change_gt_primary_sd=mean(relative_to_primary_sd>1,na.rm=TRUE))
)

scatter_row <- (prec_plots$scatter | phi_plots$scatter) + plot_layout(guides="collect")
change_row <- prec_plots$change | phi_plots$change

final_prior_plots <- (scatter_row/change_row) +
  plot_annotation(
    title=paste0("Prior sensitivity for ",sensitivity_model),
    subtitle="Top row: primary-prior versus sensitivity-prior LAD means; bottom row: thirty largest LTLA changes",
    theme=theme(text=element_text(family=map_family),
                plot.title=element_text(family=map_family,size=11.5,face="bold",hjust=0),
                plot.subtitle=element_text(family=map_family,size=8.5,hjust=0))
  ) &
  theme(legend.position="bottom",legend.box="horizontal",
        legend.box.spacing=unit(0,"mm"))

print(final_prior_plots)
prior_sensitivity_summaries
```

# 18. Marginal MRP maps

The prediction grid contains the eight LTLA × age × employment Census cells. The maps below show age and employment separately; for each displayed category, posterior draws are population-averaged over the other MRP dimension before posterior summaries are calculated.

```{r marginal-maps, fig.width=8.27, fig.height=5.83, dpi=300}
districts_plot <- districts |> select(LAD23CD) |> st_make_valid() |> st_simplify(dTolerance = 100)

mrp_display_vars <- c("agelong","remploya")
var_labels <- c(agelong="age group",remploya="employment status")
map_family <- "roboto_slab"

pretty_category <- function(x) {
  recode(as.character(x),
    "age_16_24"="16–24","age_25_34"="25–34","age_35_64"="35–64","age_65_plus"="65+",
    "employed"="Employed","unemployed_or_economically_inactive"="Unemployed or economically inactive",
    .default=as.character(x)
  ) |> stringr::str_wrap(width=17)
}

marginalise_cell_draws <- function(prediction_result, prediction_grid, display_var, model_name) {
  draws <- prediction_result$cell_draws
  stopifnot(nrow(draws)==nrow(prediction_grid), display_var %in% names(prediction_grid))
  key <- interaction(prediction_grid$LAD23CD, prediction_grid[[display_var]], drop=TRUE, lex.order=TRUE)
  idx <- split(seq_len(nrow(prediction_grid)), key)

  marginal_draws <- do.call(rbind, lapply(idx, function(i) {
    w <- prediction_grid$N[i] / sum(prediction_grid$N[i])
    colSums(draws[i,,drop=FALSE] * w)
  }))

  meta <- bind_rows(lapply(idx, function(i) prediction_grid[i[1], c("LAD23CD",display_var), drop=FALSE]))
  bind_cols(meta, summarise_draws(marginal_draws)) |>
    mutate(ci_width=upper-lower, model=model_name, displayed_variable=display_var)
}

marginal_predictions <- imap_dfr(predictions, \(p,m)
  map_dfr(mrp_display_vars, \(v) marginalise_cell_draws(p, prediction_data, v, m))) |>
  mutate(category=coalesce(as.character(agelong),as.character(remploya)))

ci_min <- 0.20
ci_max <- max(marginal_predictions$ci_width, na.rm=TRUE)
ci_breaks <- seq(ci_min, ci_max, length.out=11)

base_map_theme <- theme_void(base_family=map_family) +
  theme(
    text=element_text(family=map_family,size=11),
    plot.title=element_text(size=11.5,face="bold",hjust=.5,margin=margin(b=1)),
    plot.subtitle=element_text(size=9,hjust=.5,lineheight=.95,margin=margin(b=2)),
    plot.title.position="plot",
    strip.text=element_text(size=9,face="bold",hjust=.5,lineheight=.9,margin=margin(t=1,b=1)),
    legend.position="bottom", legend.direction="horizontal",
    legend.title=element_text(size=9.5,face="bold"),
    legend.text=element_text(size=8.5),
    legend.margin=margin(t=-2), legend.box.margin=margin(t=-3),
    panel.spacing=unit(2,"mm"), plot.margin=margin(2,3,2,3)
  )

make_marginal_maps <- function(prediction, model_name, display_var) {
  plot_data <- prediction |>
    filter(model==model_name, displayed_variable==display_var) |>
    mutate(category=factor(pretty_category(category), levels=unique(pretty_category(category))))

  map_data <- districts_plot |> left_join(plot_data, by="LAD23CD", relationship="one-to-many")
  ncol_facets <- if (n_distinct(plot_data$category)<=2) n_distinct(plot_data$category) else 2L

  mean_breaks <- seq(0.40, 0.60, by=0.02)
  if (length(mean_breaks)<3) mean_breaks <- pretty(range(plot_data$mean,na.rm=TRUE), n=10)

  mean_cols <- c("#543005","#8c510a","#bf812d","#dfc27d","#f6e8c3","#c7eae5","#80cdc1","#35978f","#01665e","#003c30")
  subtitle_txt <- stringr::str_wrap("Results are population-averaged over the other MRP dimension",width=52)

  mean_map <- ggplot(map_data) +
    geom_sf(aes(fill=mean), colour="white", linewidth=.02) +
    facet_wrap(~category, ncol=ncol_facets) +
    scale_fill_stepsn(
      colours=mean_cols,breaks=mean_breaks,limits=c(0.40,0.60), labels=label_percent(accuracy=1), name="Predicted reporting", oob=scales::squish,
      guide=guide_coloursteps(
        direction="horizontal", title.position="top", title.hjust=.5,
        barwidth=unit(72,"mm"), barheight=unit(4.5,"mm"),
        ticks=TRUE, even.steps=TRUE
      )
    ) +
    coord_sf(datum=NA, expand=FALSE) +
    labs(title=paste0(model_name," – reporting probability by ",var_labels[[display_var]]), subtitle=subtitle_txt) +
    base_map_theme
  
  uncertainty_map <- ggplot(map_data) +
    geom_sf(aes(fill=ci_width), colour="white", linewidth=.02) +
    facet_wrap(~category, ncol=ncol_facets) +
    scale_fill_stepsn(
      colours = c("#fcfbfd", "#efe7f2", "#d0b8dd", "#b07cc6", "#8c4fb0", "#6a2c91", "#54278f", "#3f007d", "#2b0057", "#14002b"),
      breaks=ci_breaks,limits=c(ci_min,ci_max),labels=label_percent(accuracy=1),name="95% CrI width", oob=scales::squish,
      guide=guide_coloursteps(
        direction="horizontal", title.position="top", title.hjust=.5,
        barwidth=unit(72,"mm"), barheight=unit(4.5,"mm"),
        ticks=TRUE, even.steps=TRUE
      )
    ) +
    coord_sf(datum=NA, expand=FALSE) +
    labs(title=paste0(model_name," – uncertainty by ",var_labels[[display_var]]),
         subtitle=subtitle_txt) +
    base_map_theme

  mean_map + uncertainty_map + plot_layout(widths=c(1,1))
}

# Folder for PNG exports
png_dir <- file.path(output_dir, "marginal_maps_png")
dir.create(png_dir, recursive = TRUE, showWarnings = FALSE)

# Build, print, and save all marginal map figures
marginal_map_plots <- list()

for (m in names(fits)) {
  vars_to_plot <- mrp_display_vars
  
  for (v in vars_to_plot) {
    p <- make_marginal_maps(marginal_predictions, m, v)
    
    file_stub <- paste(m, v, sep = "_") |>
      stringr::str_replace_all("[^A-Za-z0-9_]+", "_")
    
    marginal_map_plots[[file_stub]] <- p
    print(p)
    
    ggsave(
      filename = file.path(png_dir, paste0(file_stub, ".png")),
      plot = p,
      width = 8.27, height = 5.83, units = "in", dpi = 300,
      bg = "white"
    )
  }
}
```



```{r overall-maps, fig.width=8.27, fig.height=5.83, dpi=300}
mean_min <- min(district_predictions$mean,na.rm=TRUE); mean_max <- max(district_predictions$mean,na.rm=TRUE)
mean_breaks <- seq(mean_min,mean_max,length.out=11)
ci_min <- min(district_predictions$ci_width,na.rm=TRUE); ci_max <- max(district_predictions$ci_width,na.rm=TRUE)
ci_breaks <- seq(ci_min,ci_max,length.out=11)

mean_cols <- c("#543005","#8c510a","#bf812d","#dfc27d","#f6e8c3",
               "#c7eae5","#80cdc1","#35978f","#01665e","#003c30")
unc_cols <- c("#fcfbfd","#efe7f2","#d0b8dd","#b07cc6","#8c4fb0",
              "#6a2c91","#54278f","#3f007d","#2b0057","#14002b")

base_map_theme <- theme_void(base_family=map_family) +
  theme(
    text=element_text(family=map_family,size=11),
    plot.title=element_text(size=11.5,face="bold",hjust=.5,margin=margin(b=1)),
    plot.subtitle=element_text(size=9,hjust=.5,lineheight=.95,margin=margin(b=2)),
    plot.title.position="plot",
    legend.position="bottom", legend.direction="horizontal",
    legend.title=element_text(size=9.5,face="bold"), legend.text=element_text(size=8.5),
    legend.margin=margin(t=-2), legend.box.margin=margin(t=-3),
    plot.margin=margin(2,3,2,3)
  )

make_overall_maps <- function(prediction,model_name) {
  map_data <- districts_plot |> left_join(prediction,by="LAD23CD")
  subtitle_txt <- stringr::str_wrap(
    "Overall LTLA estimates after poststratification across age × employment cells",
    width=52
  )

  mean_map <- ggplot(map_data) +
    geom_sf(aes(fill=mean),colour="white",linewidth=.03) +
    scale_fill_stepsn(
      colours=mean_cols, breaks=mean_breaks, limits=c(mean_min,mean_max),
      labels=label_percent(accuracy=1), name="Predicted reporting", oob=scales::squish,
      guide=guide_coloursteps(direction="horizontal",title.position="top",title.hjust=.5,
                              barwidth=unit(72,"mm"),barheight=unit(4.5,"mm"),
                              ticks=TRUE,even.steps=TRUE)
    ) +
    coord_sf(datum=NA,expand=FALSE) +
    labs(title=paste0(model_name," – overall LTLA reporting probability"),subtitle=subtitle_txt) +
    base_map_theme

  uncertainty_map <- ggplot(map_data) +
    geom_sf(aes(fill=ci_width),colour="white",linewidth=.03) +
    scale_fill_stepsn(
      colours=unc_cols, breaks=ci_breaks, limits=c(ci_min,ci_max),
      labels=label_percent(accuracy=1), name="95% CrI width", oob=scales::squish,
      guide=guide_coloursteps(direction="horizontal",title.position="top",title.hjust=.5,
                              barwidth=unit(72,"mm"),barheight=unit(4.5,"mm"),
                              ticks=TRUE,even.steps=TRUE)
    ) +
    coord_sf(datum=NA,expand=FALSE) +
    labs(title=paste0(model_name," – overall LTLA uncertainty"),subtitle=subtitle_txt) +
    base_map_theme

  mean_map + uncertainty_map + plot_layout(widths=c(1,1))
}

for (m in names(fits))
  print(make_overall_maps(filter(district_predictions,model==m),m))
```
```{r getis-ord-maps, fig.width=8.27, fig.height=5.83, dpi=300}
# Gi* requires the focal LTLA to be included in its own neighbourhood
nb_star <- include.self(nb)
listw_star <- nb2listw(nb_star,style="B",zero.policy=TRUE)

gi_cols <- c(
  "High reporting – high certainty"="#b2182b",
  "High reporting – low certainty"="#ef8a62",
  "Low reporting – high certainty"="#2166ac",
  "Low reporting – low certainty"="#67a9cf",
  "No significant clustering"="#dddddd"
)

gi_levels <- names(gi_cols)

getis_ord_results <- map_dfr(names(fits),\(m) {
  x <- district_predictions |>
    filter(model==m) |>
    select(LAD23CD,mean,ci_width) |>
    right_join(
      tibble(LAD23CD=graph_levels,spatial_id=seq_along(graph_levels)),
      by="LAD23CD"
    ) |>
    arrange(spatial_id)

  stopifnot(!anyNA(x$mean),nrow(x)==length(nb_star))

  x |>
    mutate(
      gi_z=as.numeric(localG(mean,listw_star,zero.policy=TRUE)),
      certainty_cut=median(ci_width,na.rm=TRUE),
      certainty=if_else(ci_width<=certainty_cut,"high","low"),
      gi_class=case_when(
        gi_z>=1.96 & certainty=="high" ~ "High reporting – high certainty",
        gi_z>=1.96 & certainty=="low"  ~ "High reporting – low certainty",
        gi_z<=-1.96 & certainty=="high"~ "Low reporting – high certainty",
        gi_z<=-1.96 & certainty=="low" ~ "Low reporting – low certainty",
        TRUE ~ "No significant clustering"
      ),
      gi_class=factor(gi_class,levels=gi_levels),
      model=m
    )
})

getis_map_theme <- theme_void(base_family=map_family) +
  theme(
    text=element_text(family=map_family),
    plot.title=element_text(size=12,face="bold",hjust=.5,margin=margin(b=2)),
    plot.subtitle=element_text(size=9,hjust=.5,lineheight=.95,margin=margin(b=3)),
    legend.position="bottom",
    legend.title=element_blank(),
    legend.text=element_text(size=8),
    legend.key.height=unit(4,"mm"),
    legend.key.width=unit(7,"mm"),
    legend.margin=margin(t=-2),
    plot.margin=margin(2,3,2,3)
  )

make_getis_map <- function(m) {
  map_data <- districts_plot |>
    left_join(filter(getis_ord_results,model==m),by="LAD23CD")

  ggplot(map_data) +
    geom_sf(aes(fill=gi_class),colour="white",linewidth=.04) +
    scale_fill_manual(values=gi_cols,drop=FALSE) +
    coord_sf(datum=NA,expand=FALSE) +
    labs(
      title=paste0(m," – local clusters of overall reporting probability"),
      subtitle="Getis–Ord Gi*: |z| ≥ 1.96; certainty based on model-specific median 95% CrI width"
    ) +
    guides(fill=guide_legend(nrow=2,byrow=TRUE)) +
    getis_map_theme
}

for(m in names(fits)) print(make_getis_map(m))

# write_csv(
#   getis_ord_results |> select(LAD23CD,model,mean,ci_width,gi_z,certainty,gi_class),
#   file.path(output_dir,"Getis_Ord_overall_reporting_by_model.csv")
# )
```

```{r}
# -----------------------------
# Global Moran's I by model
# -----------------------------
listw <- nb2listw(nb,style="W",zero.policy=TRUE)

moran_results <- map_dfr(names(fits), \(m) {
  x <- district_predictions |>
    filter(model==m) |>
    right_join(tibble(LAD23CD=graph_levels,spatial_id=seq_along(graph_levels)),
               by="LAD23CD") |>
    arrange(spatial_id)

  mt <- moran.test(x$mean,listw,zero.policy=TRUE)

  tibble(
    model=m,
    morans_i=unname(mt$estimate["Moran I statistic"]),
    expected_i=unname(mt$estimate["Expectation"]),
    p_value=mt$p.value
  )
})

moran_results
```

```{r}
library(showtext)
showtext_auto()
map_family <- "roboto_slab"
```


```{r overall-maps-all-models-a5, fig.width=8.27, fig.height=6.7, dpi=300}
mean_min <- min(district_predictions$mean,na.rm=TRUE); mean_max <- max(district_predictions$mean,na.rm=TRUE)
ci_min <- min(district_predictions$ci_width,na.rm=TRUE); ci_max <- max(district_predictions$ci_width,na.rm=TRUE)
mean_breaks <- seq(mean_min,mean_max,length.out=11); ci_breaks <- seq(ci_min,ci_max,length.out=11)

mean_cols <- c("#543005","#8c510a","#bf812d","#dfc27d","#f6e8c3","#c7eae5","#80cdc1","#35978f","#01665e","#003c30")
unc_cols <- c("#fcfbfd","#efe7f2","#d0b8dd","#b07cc6","#8c4fb0","#6a2c91","#54278f","#3f007d","#2b0057","#14002b")

map_all <- districts_plot |>
  left_join(district_predictions |> select(LAD23CD,model,mean,ci_width), by="LAD23CD", relationship="one-to-many") |>
  mutate(model=factor(model,levels=names(fits)), block=if_else(as.integer(model)<=3,"Models 1–3","Models 4–6"))

map_theme <- theme_void(base_family=map_family) +
  theme(
    text=element_text(family=map_family), aspect.ratio=1,
    plot.title=element_text(size=9.5,face="bold",hjust=.5,margin=margin(b=1)),
    strip.background=element_blank(), strip.text.x=element_text(size=8.5,margin=margin(b=1)),
    panel.spacing.y=unit(.3,"mm"), plot.margin=margin(.2,.1,.2,.1)
  )

mean_scale <- scale_fill_stepsn(
  colours=mean_cols, breaks=mean_breaks, limits=c(mean_min,mean_max), labels=label_percent(accuracy=1),
  name="Predicted\nreporting", oob=scales::squish,
  guide=guide_coloursteps(direction="vertical", title.position="top", barwidth=unit(4,"mm"), barheight=unit(34,"mm"),
                          ticks=TRUE, even.steps=TRUE)
)

unc_scale <- scale_fill_stepsn(
  colours=unc_cols, breaks=ci_breaks, limits=c(ci_min,ci_max), labels=label_percent(accuracy=1),
  name="95% CrI\nwidth", oob=scales::squish,
  guide=guide_coloursteps(direction="vertical", title.position="top", barwidth=unit(4,"mm"), barheight=unit(34,"mm"),
                          ticks=TRUE, even.steps=TRUE)
)

make_col <- function(dat, fill_var, title_txt=NULL, scale_obj) {
  ggplot(dat) + geom_sf(aes(fill=.data[[fill_var]]), colour="white", linewidth=.015) +
    facet_wrap(~model, ncol=1, strip.position="top") + scale_obj +
    coord_sf(datum=NA, expand=FALSE) + labs(title=title_txt) + map_theme
}

m13b <- make_col(filter(map_all,block=="Models 1–3"), "mean", "Reporting probability", mean_scale)
m46b <- make_col(filter(map_all,block=="Models 4–6"), "mean", NULL, mean_scale)
u13b <- make_col(filter(map_all,block=="Models 1–3"), "ci_width", "Uncertainty", unc_scale)
u46b <- make_col(filter(map_all,block=="Models 4–6"), "ci_width", NULL, unc_scale)

m13 <- m13b + theme(legend.position="none", plot.margin=margin(.2,-20,.2,.1))
m46 <- m46b + theme(legend.position="none", plot.margin=margin(.2,.1,.2,-20))
u13 <- u13b + theme(legend.position="none", plot.margin=margin(.2,-2,.2,.1))
u46 <- u46b + theme(legend.position="none", plot.margin=margin(.2,.1,.2,-2))

mean_key <- wrap_elements(cowplot::get_legend(
  m13b + theme(legend.position="right",
               legend.title=element_text(family=map_family,size=8),
               legend.text=element_text(family=map_family,size=7))
)) & theme(plot.margin=margin(10,20,10,20))

unc_key <- wrap_elements(cowplot::get_legend(
  u13b + theme(legend.position="right",
               legend.title=element_text(family=map_family,size=8),
               legend.text=element_text(family=map_family,size=7))
)) & theme(plot.margin=margin(10,20,10,20))

final_overall_maps <- (
  m13 + m46 + plot_spacer() + mean_key +
  u13 + u46 + plot_spacer() + unc_key +
    plot_layout(widths=c(1,1,.10,.24, 1,1,.10,.24)) +
    plot_annotation(
      title="Overall LTLA reporting estimates across models",
      subtitle="Posterior reporting probability (left block) and corresponding 95% credible-interval width (right block)",
      theme=theme(
        plot.title=element_text(family=map_family,size=10.5,face="bold",hjust=.5,margin=margin(t=12,b=2)),
        plot.subtitle=element_text(family=map_family,size=8.5,hjust=.5,margin=margin(b=14))
      )
    )
) & theme(plot.margin=margin(0,0,0,0))

final_overall_maps
```
```{r}
install.packages("shadowtext")
library(shadowtext)
```

```{r direct-gor-maps, fig.width=5.27, fig.height=3.83, dpi=300}
gor_plot <- districts |>
  st_make_valid() |>
  mutate(gor = as.character(gor)) |>
  group_by(gor) |>
  summarise(.groups = "drop") |>
  st_make_valid() |>
  st_simplify(dTolerance = 1000)

direct_gor_map <- direct_gor |>
  mutate(
    gor = as.character(gor),
    ci_width = upper - lower,
    estimate_lab = scales::percent(estimate, accuracy = 0.1),
    ci_lab = scales::percent(ci_width, accuracy = 0.1)
  )

map_data <- gor_plot |>
  left_join(direct_gor_map, by = "gor")

label_points <- map_data |> st_point_on_surface()
xy <- st_coordinates(label_points)

label_data <- label_points |>
  mutate(x = xy[,1], y = xy[,2]) |>
  st_drop_geometry() |>
  mutate(
    x = if_else(gor == "london", x + 30000, x)
  )

mean_min <- min(map_data$estimate, na.rm = TRUE); mean_max <- max(map_data$estimate, na.rm = TRUE)
ci_min <- min(map_data$ci_width, na.rm = TRUE); ci_max <- max(map_data$ci_width, na.rm = TRUE)
mean_breaks <- seq(mean_min, mean_max, length.out = 11)
ci_breaks <- seq(ci_min, ci_max, length.out = 11)

mean_cols <- c("#543005","#8c510a","#bf812d","#dfc27d","#f6e8c3","#c7eae5","#80cdc1","#35978f","#01665e","#003c30")
unc_cols  <- c("#fcfbfd","#efe7f2","#d0b8dd","#b07cc6","#8c4fb0","#6a2c91","#54278f","#3f007d","#2b0057","#14002b")

base_map_theme <- theme_void(base_family = map_family) +
  theme(
    text = element_text(family = map_family, size = 11),
    plot.title = element_text(family = map_family, size = 11.5, face = "bold", hjust = .5, margin = margin(b = 1)),
    plot.subtitle = element_text(family = map_family, size = 9, hjust = .5, lineheight = .95, margin = margin(b = 2)),
    plot.title.position = "plot",
    plot.margin = margin(2, 3, 2, 3)
  )

# subtitle_txt <- stringr::str_wrap(
#   "Direct survey-weighted reporting estimates by Government Office Region",
#   width = 52
# )

p_direct_mean <- ggplot(map_data) +
  geom_sf(aes(fill = estimate), colour = "white", linewidth = .25) +
  shadowtext::geom_shadowtext(
    data = label_data,
    aes(x = x, y = y, label = estimate_lab),
    family = map_family, size = 3.2, lineheight = .9,
    colour = "black", bg.colour = "white", bg.r = 0.15
  ) +
  scale_fill_stepsn(
    colours = mean_cols, breaks = mean_breaks, limits = c(mean_min, mean_max),
    oob = scales::squish
  ) +
  coord_sf(datum = NA, expand = FALSE) +
  labs(title = "Direct GOR reporting estimate") +
  base_map_theme +
  theme(legend.position = "none")

p_direct_unc <- ggplot(map_data) +
  geom_sf(aes(fill = ci_width), colour = "white", linewidth = .25) +
  shadowtext::geom_shadowtext(
    data = label_data,
    aes(x = x, y = y, label = ci_lab),
    family = map_family, size = 3.2, lineheight = .9,
    colour = "black", bg.colour = "white", bg.r = 0.15
  ) +
  scale_fill_stepsn(
    colours = unc_cols, breaks = ci_breaks, limits = c(ci_min, ci_max),
    oob = scales::squish
  ) +
  coord_sf(datum = NA, expand = FALSE) +
  labs(title = "Direct GOR uncertainty") +
  base_map_theme +
  theme(legend.position = "none")

p_direct_mean + p_direct_unc + plot_layout(widths = c(1, 1))
```

```{r disaggregation-spatial-levels, fig.width=8.27, fig.height=5.83, dpi=300}
# -----------------------------
# GOR → LTLA spatial levels
# -----------------------------
gor_plot <- districts |> group_by(gor) |> summarise(.groups = "drop")
bb <- st_bbox(districts)

outline_theme <- theme_void(base_family = map_family) +
  theme(
    text = element_text(family = map_family),
    plot.title = element_text(size = 12, face = "bold", hjust = .5),
    plot.subtitle = element_text(size = 9, hjust = .5),
    plot.margin = margin(2, 2, 2, 2)
  )

fixed_sf <- coord_sf(
  xlim = c(bb["xmin"], bb["xmax"]),
  ylim = c(bb["ymin"], bb["ymax"]),
  datum = NA,
  expand = FALSE
)

p_gor <- ggplot() +
  geom_sf(data = gor_plot, fill = "white", colour = "grey25", linewidth = .25) +
  fixed_sf +
  labs(title = "GOR", subtitle = "Observed spatial level") +
  outline_theme

p_ltla <- ggplot() +
  geom_sf(data = districts, fill = "white", colour = "grey25", linewidth = .25) +
  #geom_sf(data = gor_plot, fill = NA, colour = "grey20", linewidth = .35) +
  fixed_sf +
  labs(title = "LTLA", subtitle = "Estimated spatial level") +
  outline_theme

p_gor | p_ltla
```
```{r}
# ------------------------------------------------------------
# GVIF for all six model specification
library(car)

offence_vars <- c(
  "vawg_official_codes","vawg_sex_force_threats",
  "vawg_domestic","vawg_harasm"
)

individual_vars <- c(
  "relig3","rnssec3","ethgrp2a","polatt7",
  "educint","nchil2","illharmONS2","nslivarr"
)

area_vars <- c("rural3","inner","imd","imd_crime")

model_terms <- list(
  "Model 1" = c("rural3","imd_crime"),
  "Model 2" = c("rural3","imd_crime",individual_vars),
  "Model 3" = area_vars,
  "Model 4" = c("rural3","imd_crime",offence_vars),
  "Model 5" = c(area_vars,individual_vars,offence_vars),
  "Model 6" = c(area_vars,individual_vars)
) |>
  map(~ c("mrp_strata", .x))

get_gvif <- function(vars, model_name) {

  d <- person_data |>
    select(police_known_all, all_of(vars)) |>
    mutate(across(all_of(vars), factor))

  fit <- glm(
    reformulate(vars, response = "police_known_all"),
    data = d,
    family = binomial()
  )

  v <- car::vif(fit)

  as.data.frame(v) |>
    rownames_to_column("variable") |>
    mutate(
      model = model_name,
      adjusted_GVIF = GVIF^(1 / (2 * Df)),
      .before = 1
    ) |>
    select(model, variable, GVIF, Df, adjusted_GVIF)
}

gvif_results <- imap_dfr(model_terms, ~ get_gvif(.x, .y))

gvif_results |>
  arrange(model, desc(adjusted_GVIF))
```


# 19. Reproducibility record

```{r session-info}
writeLines(capture.output(sessionInfo()), file.path(output_dir, "sessionInfo.txt"))
sessionInfo()
```
